PANCHANGE_data_preprocessing

### 2020-08-11 PANCHANGE_data_preprocessing2020-07-15.Rmd

@@

prepandemic
```{r phq prepan missingness}
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    na_per_person_phq_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[phq.items_prepan])
      )
  )

prepan.phq.raw.id %>%
  freq(na_per_person_phq_prepan,
       cumul = F)
```

```{r phq prepan missingness 8 items}
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    phq.missing_only_suicide_item_prepan = 
      if_else(
        condition = na_per_person_phq_prepan == 1 & 
          is.na(phq.dead_hurting_thoughts_prepan_numeric), 
        true = 1,
        false = 0,
        missing = NA_real_
      )
  )

prepan.phq.raw.id %>%
  freq(phq.missing_only_suicide_item_prepan,
       cumul = F)
```

all times
```{r phq prepan sum score if missigng items set to NA - all items}
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    phq.sum_score_prepan =
      if_else(condition = na_per_person_phq_prepan > 0,
              true = NA_real_,
              false = phq.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

@@

prepan.phq.raw.id %>%
  descr(phq.sum_score_prepan, stats = "common")
```

@@

8 items
```{r phq prepan sum score if missigng items set to NA - 8 items}
glad.phq.raw.id <- glad.phq.raw.id %>%
  mutate(
    phq.sum_score_8items_prepan =
      if_else(condition = na_per_person_phq_prepan > 1 &
                phq.missing_only_suicide_item_prepan == 1,
              true = NA_real_,
              false = phq.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

glad.phq.raw.id %>%
  descr(phq.sum_score_8items_prepan, stats = "common")

glad.phq.raw.id %>%
  descr(phq.sum_score_8items_prepan, stats = "common")
```

```{r glad phq head}
head(glad.phq.raw.id)
```

```{r glad phq summary}
summarytools::dfSummary(
  glad.phq.raw.id,
  graph.col = F,
  valid.col = F
  )
```

@@

```{r glad ocir cols}
glad.ocir.raw <- readRDS(file = paste0(data.raw_path, "/glad/ocir_glad.rds"))
dim(glad.ocir.raw)
colnames(glad.ocir.raw)
```

```{r glad ocir cleaning}
glad.ocir.raw.id <- glad.ocir.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
    ocir.arranged_things_change_upset_prepan = ocir.arranged_things_change_upset,
    ocir.repeat_feel_numbers_prepan = ocir.repeat_feel_numbers,
    ocir.clean_simply_wash_feel_prepan = ocir.clean_simply_wash_feel,
    ocir.mind_upset_unpleasant_thoughts_prepan = ocir.mind_upset_unpleasant_thoughts,
    ocir.afraid_avoid_throwing_things_prepan = ocir.afraid_avoid_throwing_things,
    ocir.light_switches_water_taps_prepan = ocir.light_switches_water_taps,
    ocir.arranged_things_prepan = ocir.arranged_things,
    ocir.good_feel_bad_numbers_prepan = ocir.good_feel_bad_numbers,
    ocir.hands_longer_wash_prepan = ocir.hands_longer_wash,
    ocir.saved_things_prepan = ocir.saved_things,
    ocir.frequently_difficulty_rid_nasty_prepan = ocir.frequently_difficulty_rid_nasty,
    ocir.i_check_things_more_often_than_necessary_prepan = ocir.i_check_things_more_often_than_necessary.,
    ocir.arranged_properly_objects_upset_prepan = ocir.arranged_properly_objects_upset,
    ocir.feel_compelled_count_things_prepan = ocir.feel_compelled_count_things,
    ocir.strangers_touch_touched_difficult_prepan = ocir.strangers_touch_touched_difficult,
    ocir.control_difficult_find_thoughts_prepan = ocir.control_difficult_find_thoughts,
    ocir.i_collect_things_i_dont_need_prepan = ocir.i_collect_things_i_dont_need
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.ocir.raw.id)
# Inspect colnames
colnames(glad.ocir.raw.id)
#Differences
dim(glad.ocir.raw)[1]-dim(glad.ocir.raw.id)[1]
```

@@

*Note that scores with missingness will be changed to NA later in the script
```{r ocir sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_prepan <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

@@

# Calulate sum score
ocir.scored_items_prepan <- scoreItems(keys = ocir.items_key_prepan,
                                        items = glad.ocir.raw.id[ocir.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 4)
# Add in column to tibble
glad.ocir.raw.id$ocir.sum_score_prepan <- ocir.scored_items_prepan$scores

# Look at values
glad.ocir.raw.id %>%
  descr(ocir.sum_score_prepan,
        stats = "common"
        )
```

prepandemic
```{r ocir prepan missingness}
glad.ocir.raw.id <- glad.ocir.raw.id %>%
  mutate(
    na_per_person_ocir_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[ocir.items_prepan])
      )
  )

glad.ocir.raw.id %>%
  freq(na_per_person_ocir_prepan,
       cumul = F)
```

```{r glad ocir head}
head(glad.ocir.raw.id)
```

```{r glad ocir summary}
summarytools::dfSummary(
  glad.ocir.raw.id,
  graph.col = F,
  valid.col = F
  )
```

>>>>

# eclude cols

```{r cols to exclude}
exclude_cols <- c("Sample",
                  "ID")
```

@@

prepandemic
```{r phq prepan missingness}
if(GLAD == TRUE | EDGI == TRUE) {
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    na_per_person_phq_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[phq.items_prepan])
      )
  )

prepan.phq.raw.id %>%
  freq(na_per_person_phq_prepan,
       cumul = F)
}
```

```{r phq prepan missingness 8 items}
if(GLAD == TRUE | EDGI == TRUE) {
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    phq.missing_only_suicide_item_prepan = 
      if_else(
        condition = na_per_person_phq_prepan == 1 & 
          is.na(phq.dead_hurting_thoughts_prepan_numeric), 
        true = 1,
        false = 0,
        missing = NA_real_
      )
  )

prepan.phq.raw.id %>%
  freq(phq.missing_only_suicide_item_prepan,
       cumul = F)
}
```

all times
```{r phq prepan sum score if missigng items set to NA - all items}
if(GLAD == TRUE | EDGI == TRUE) {
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    phq.sum_score_prepan =
      if_else(condition = na_per_person_phq_prepan > 0,
              true = NA_real_,
              false = phq.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

@@

prepan.phq.raw.id %>%
  descr(phq.sum_score_prepan, stats = "common")
}
```

@@

8 items
```{r phq prepan sum score if missigng items set to NA - 8 items}
if(GLAD == TRUE | EDGI == TRUE) {
prepan.phq.raw.id <- prepan.phq.raw.id %>%
  mutate(
    phq.sum_score_8items_prepan =
      if_else(condition = na_per_person_phq_prepan > 1 &
                phq.missing_only_suicide_item_prepan == 1,
              true = NA_real_,
              false = phq.sum_score_prepan_raw,
              missing = NA_real_
      )
  )

prepan.phq.raw.id %>%
  descr(phq.sum_score_8items_prepan, stats = "common")

prepan.phq.raw.id %>%
  descr(phq.sum_score_8items_prepan, stats = "common")
}
```

```{r glad phq head}
if(GLAD == TRUE | EDGI == TRUE) {
head(prepan.phq.raw.id)
}
```

```{r glad phq summary}
if(GLAD == TRUE | EDGI == TRUE) {
summarytools::dfSummary(
  prepan.phq.raw.id,
  graph.col = F,
  valid.col = F
  )
}
```

@@

```{r glad ocir cols}
if(GLAD == TRUE){
glad.ocir.raw <- readRDS(file = paste0(data.raw_path, "/glad/ocir_glad.rds"))
dim(glad.ocir.raw)
colnames(glad.ocir.raw)
}
```

```{r glad ocir cleaning}
if(GLAD == TRUE){
glad.ocir.raw.id <- glad.ocir.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
    ocir.arranged_things_change_upset_prepan = ocir.arranged_things_change_upset,
    ocir.repeat_feel_numbers_prepan = ocir.repeat_feel_numbers,
    ocir.clean_simply_wash_feel_prepan = ocir.clean_simply_wash_feel,
    ocir.mind_upset_unpleasant_thoughts_prepan = ocir.mind_upset_unpleasant_thoughts,
    ocir.afraid_avoid_throwing_things_prepan = ocir.afraid_avoid_throwing_things,
    ocir.light_switches_water_taps_prepan = ocir.light_switches_water_taps,
    ocir.arranged_things_prepan = ocir.arranged_things,
    ocir.good_feel_bad_numbers_prepan = ocir.good_feel_bad_numbers,
    ocir.hands_longer_wash_prepan = ocir.hands_longer_wash,
    ocir.saved_things_prepan = ocir.saved_things,
    ocir.frequently_difficulty_rid_nasty_prepan = ocir.frequently_difficulty_rid_nasty,
    ocir.i_check_things_more_often_than_necessary_prepan = ocir.i_check_things_more_often_than_necessary.,
    ocir.arranged_properly_objects_upset_prepan = ocir.arranged_properly_objects_upset,
    ocir.feel_compelled_count_things_prepan = ocir.feel_compelled_count_things,
    ocir.strangers_touch_touched_difficult_prepan = ocir.strangers_touch_touched_difficult,
    ocir.control_difficult_find_thoughts_prepan = ocir.control_difficult_find_thoughts,
    ocir.i_collect_things_i_dont_need_prepan = ocir.i_collect_things_i_dont_need
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.ocir.raw.id)
# Inspect colnames
colnames(glad.ocir.raw.id)
#Differences
dim(glad.ocir.raw)[1]-dim(glad.ocir.raw.id)[1]
}
```

```{r glad ocir prepan copy}
if(GLAD == TRUE) {
  prepan.ocir.raw.id <- glad.ocir.raw.id
}
```

# edgi ocir

```{r edgi ocir cols}
if(EDGI == TRUE){
  edgi.ocir.raw <- readRDS(file = paste0(data.raw_path, "/edgi/ocir_edgi.rds"))
  dim(edgi.ocir.raw)
  colnames(edgi.ocir.raw)
}
```

```{r edgi ocir cleaning}
if(EDGI == TRUE){
  edgi.ocir.raw.id <- edgi.ocir.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.prepandemic = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
      ocir.arranged_things_change_upset_prepan = ocir.arranged_things_change_upset,
      ocir.repeat_feel_numbers_prepan = ocir.repeat_feel_numbers,
      ocir.clean_simply_wash_feel_prepan = ocir.clean_simply_wash_feel,
      ocir.mind_upset_unpleasant_thoughts_prepan = ocir.mind_upset_unpleasant_thoughts,
      ocir.afraid_avoid_throwing_things_prepan = ocir.afraid_avoid_throwing_things,
      ocir.light_switches_water_taps_prepan = ocir.light_switches_water_taps,
      ocir.arranged_things_prepan = ocir.arranged_things,
      ocir.good_feel_bad_numbers_prepan = ocir.good_feel_bad_numbers,
      ocir.hands_longer_wash_prepan = ocir.hands_longer_wash,
      ocir.saved_things_prepan = ocir.saved_things,
      ocir.frequently_difficulty_rid_nasty_prepan = ocir.frequently_difficulty_rid_nasty,
      ocir.i_check_things_more_often_than_necessary_prepan = ocir.i_check_things_more_often_than_necessary.,
      ocir.arranged_properly_objects_upset_prepan = ocir.arranged_properly_objects_upset,
      ocir.feel_compelled_count_things_prepan = ocir.feel_compelled_count_things,
      ocir.strangers_touch_touched_difficult_prepan = ocir.strangers_touch_touched_difficult,
      ocir.control_difficult_find_thoughts_prepan = ocir.control_difficult_find_thoughts,
      ocir.i_collect_things_i_dont_need_prepan = ocir.i_collect_things_i_dont_need
    ) %>%
    add_numeric(., exclude = exclude_cols) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(edgi.ocir.raw.id)
  # Inspect colnames
  colnames(edgi.ocir.raw.id)
  #Differences
  dim(edgi.ocir.raw)[1]-dim(edgi.ocir.raw.id)[1]
}
```

```{r edgi ocir prepan copy}
if(EDGI == TRUE) {
  prepan.ocir.raw.id <- edgi.ocir.raw.id
}
```

@@

*Note that scores with missingness will be changed to NA later in the script
```{r ocir sum score prepan}
if(GLAD == TRUE | EDGI == TRUE) {
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_prepan <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

@@

# Calulate sum score
ocir.scored_items_prepan <- scoreItems(keys = ocir.items_key_prepan,
                                        items = prepan.ocir.raw.id[ocir.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 4)
# Add in column to tibble
prepan.ocir.raw.id$ocir.sum_score_prepan <- ocir.scored_items_prepan$scores

# Look at values
prepan.ocir.raw.id %>%
  descr(ocir.sum_score_prepan,
        stats = "common"
        )
}
```

prepandemic
```{r ocir prepan missingness}
if(GLAD == TRUE | EDGI == TRUE) {
prepan.ocir.raw.id <- prepan.ocir.raw.id %>%
  mutate(
    na_per_person_ocir_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[ocir.items_prepan])
      )
  )

prepan.ocir.raw.id %>%
  freq(na_per_person_ocir_prepan,
       cumul = F)
}
```

```{r glad ocir head}
if(GLAD == TRUE | EDGI == TRUE) {
head(prepan.ocir.raw.id)
}
```

```{r glad ocir summary}
if(GLAD == TRUE | EDGI == TRUE) {
summarytools::dfSummary(
  prepan.ocir.raw.id,
  graph.col = F,
  valid.col = F
  )
}
```
